---
title: "01-data-cleaning"
output: html_notebook
---
## About this code 

This is first attempt for data cleaning for paper 2




## Loading required libraries

```{r}

library(MEPS)     
library(survey)
library(tidyverse)
library(haven)
library(labelled)
library(broom)
```

# Set survey option for lonely PSUs
```{r}
options(survey.lonely.psu="adjust")

```

## Loading Dataset

Load datasets and Rename year-specific variables prior to combining  
```{r}
for (year in 2011:2019) {
    
    assign(paste0("fyc", year),
    read_MEPS(year = year, type = "FYC"))  ## loading dataset
    
    
    dataset_name <- paste0("fyc", year)
    new_dataset_name <- paste0("fyc", year)
    
    assign(new_dataset_name,
           get(dataset_name) %>%
               rename(
                   PERWT  = paste0("PERWT", substr(year, 3, 4), "F"), # substr(), the function extracts a substring from the year, starting from the third 
                                                                        #character position and ending at the fourth character position
                                                                           
                   AGE = paste0("AGE", substr(year, 3, 4), "X"),
                   MARRYX = paste0("MARRY", substr(year, 3, 4), "X"),
                   POVCAT = paste0("POVCAT", substr(year, 3, 4)),
                   POVLEV = paste0("POVLEV", substr(year, 3, 4) ),
                   REGION = paste0("REGION", substr(year, 3, 4)),
                   INSURC = paste0("INSURC", substr(year, 3, 4)), #Insurance
                   MCDEV = paste0("MCDEV", substr(year, 3, 4)), 
                   MCREV = paste0("MCREV", substr(year, 3, 4)), 
                   PRVEV = paste0("PRVEV", substr(year, 3, 4)) , 
                   UNINS = paste0("UNINS", substr(year, 3, 4)) ,
                   INSCOV = paste0("INSCOV", substr(year, 3, 4)) ,
                   PRIV = paste0("PRIV", substr(year, 3, 4))  
               ))
}



```

years 2011 to 2019 are my timeline span, however I need year 2020 in order to create screening variable for one of the panel in 2019.

```{r}

fyc2020<-read_MEPS(year = 2020, type = "FYC")

```


Now, I move to subset variable I need, for each year and clean them before I can append them toghether. I might be able to shorten this code later on in the process. 

#2019

```{r}

fyc19 <- fyc2019 %>% select(
                DUPERSID,DUID,PANEL, PID,PSTATS31 , PSTATS42 , PSTATS53 , PERWT, VARSTR, VARPSU,SPOUID31, # IDS and weights
                AGE, SEX, RACETHX,HISPANX, MARRYX, FAMS1231,  EMPST31, POVCAT,POVLEV, DOBYY ,REGION,HIDEG, # Demographic
                BORNUSA, YRSINUS, OTHLGSPK,  HWELLSPK, WHTLGSPK, ADLANG42, # Immigration
                INSURC,MCDEV, MCREV, PRVEV, UNINS,INSCOV, PRIV ,#Insurance
                CABREAST, CACERVIX, CACOLON, CAPROSTA, CANCERDX, CAOTHER # cancer diagnoses
     )


```
fixing education:

```{r}
# 1: No Dgree

fyc19$HIDEG<-replace(fyc19$HIDEG, fyc19$HIDEG %in% c(2,3,7),2)# 2: GED or High school  2,3
fyc19$HIDEG<-replace(fyc19$HIDEG, fyc19$HIDEG == 4,3) # 3: Bachelors
fyc19$HIDEG<-replace(fyc19$HIDEG, fyc19$HIDEG %in% c(5,6),4) # 4: Master's and doctorate

fyc19<-rename(fyc19, EDUC=HIDEG)




```

 
#2018

```{r}

fyc18  <- fyc2018 %>% select(
    DUPERSID,DUID,PANEL, PID,PSTATS31 , PSTATS42 , PSTATS53 , PERWT, VARSTR, VARPSU,SPOUID31, # IDS and weights
    AGE, SEX, RACETHX,HISPANX, MARRYX, FAMS1231,  EMPST31, POVCAT,POVLEV, DOBYY , REGION,HIDEG, # Demographic
    BORNUSA, YRSINUS, OTHLGSPK,  HWELLSPK, WHTLGSPK, ADLANG42, # Immigration
    INSURC,MCDEV, MCREV, PRVEV, UNINS,INSCOV, PRIV ,#Insurance
    CABREAST, CACERVIX, CACOLON, CAPROSTA, CANCERDX, CAOTHER, # cancer diagnoses
    ADUTRM42, ADPAP42,ADPAPG42, #Cervical Cancer
    ADCOLN42, ADCLNS42, ADSGMD42, ADBLDS42, # Colorectal Cancer
    ADBRST42,ADMMGR42, # Breast Cancer
    ADPROS42, PSA53 # Prostate Cancer
)

```

cleaning education:
```{r}
# 1: No Dgree

fyc18$HIDEG<-replace(fyc18$HIDEG, fyc18$HIDEG %in% c(2,3,7),2)# 2: GED or High school  2,3
fyc18$HIDEG<-replace(fyc18$HIDEG, fyc18$HIDEG ==4,3) # 3: Bachelors
fyc18$HIDEG<-replace(fyc18$HIDEG, fyc18$HIDEG %in% c(5,6),4) # 4: Master's and doctorate


fyc18<-rename(fyc18, EDUC=HIDEG)


```


#2017

```{r}
fyc17  <- fyc2017 %>% select(
    DUPERSID,DUID,PANEL, PID,PSTATS31 , PSTATS42 , PSTATS53 , PERWT, VARSTR, VARPSU,SPOUID31, # IDS and weights
    AGE, SEX, RACETHX,HISPANX,MARRYX, FAMS1231,  EMPST31, POVCAT,POVLEV, DOBYY ,REGION,HIDEG, # Demographic
    BORNUSA, YRSINUS,  OTHLANG,LANGSPK, HWELLSPE, ADLANG42, # Immigration
    INSURC,MCDEV, MCREV, PRVEV, UNINS,INSCOV, PRIV ,#Insurance
    CABREAST, CACERVIX, CACOLON, CAPROSTA, CANCERDX, CAOTHER # cancer diagnoses
)

```

cleaning education:

```{r}
# 1: No Dgree

fyc17$HIDEG<-replace(fyc17$HIDEG, fyc17$HIDEG %in% c(2,3,7),2)# 2: GED or High school  2,3

fyc17$HIDEG<-replace(fyc17$HIDEG, fyc17$HIDEG ==4,3) # 3: Bachelors


fyc17$HIDEG<-replace(fyc17$HIDEG, fyc17$HIDEG %in% c(5,6),4) # 4: Master's and doctorate


table(fyc17$PANEL, fyc17$HIDEG)



fyc17<-rename(fyc17, EDUC=HIDEG)




```



#2016

```{r}

fyc16  <- fyc2016 %>% select(
    DUPERSID,DUID,PANEL, PID,PSTATS31 , PSTATS42 , PSTATS53 , PERWT, VARSTR, VARPSU,SPOUID31, # IDS and weights
    AGE, SEX, RACETHX,HISPANX,MARRYX, FAMS1231,  EMPST31, POVCAT,POVLEV, DOBYY ,REGION,HIDEG, # Demographic
    BORNUSA, YRSINUS, OTHLANG,LANGSPK, HWELLSPE, ADLANG42, # Immigration
    INSURC,MCDEV, MCREV, PRVEV, UNINS,INSCOV, PRIV ,#Insurance
    CABREAST, CACERVIX, CACOLON, CAPROSTA, CANCERDX, CAOTHER, # Cancer diagnoses
    HYSTER53, PAPSMR53, #Cervical Cancer
    CLNTST53, CLNTRE53,SGMTST53,SGMTRE53 , # Colorectal Cancer
    MAMOGR53, # Breast Cancer
    PSA53 # Prostate Cancer:
)



```
I need to make education consistent for all years
```{r}
# 1: No Dgree

fyc16$HIDEG<-replace(fyc16$HIDEG, fyc16$HIDEG %in% c(2,3,7),2)# 2: GED or High school  2,3

fyc16$HIDEG<-replace(fyc16$HIDEG, fyc16$HIDEG ==4,3) # 3: Bachelors


fyc16$HIDEG<-replace(fyc16$HIDEG, fyc16$HIDEG %in% c(5,6),4) # 4: Master's and doctorate


table(fyc16$PANEL, fyc16$HIDEG)



fyc16<-rename(fyc16, EDUC=HIDEG)




```

Now creating a cancer screening variables:

I decided to only look at colonscopy and semidoscopy and not the  BLD STOOL TST
thus I drop the BSTST53, BSTSRE53,

```{r}

CLNTST53

ADCLNS42

```


#2015

```{r}

fyc15  <- fyc2015 %>% select(
    DUPERSID,DUID,PANEL, PID,PSTATS31 , PSTATS42 , PSTATS53 , PERWT, VARSTR, VARPSU,SPOUID31, # IDS and weights
    AGE, SEX, RACETHX,HISPANX,MARRYX, FAMS1231, EMPST31, POVCAT,POVLEV, DOBYY ,REGION,EDRECODE, # Demographic
    BORNUSA, YRSINUS, OTHLANG,LANGSPK, HWELLSPE, ADLANG42, # Immigration
    INSURC,MCDEV, MCREV, PRVEV, UNINS,INSCOV, PRIV ,#Insurance
    CABREAST, CACERVIX, CACOLON, CAPROSTA, CANCERDX, CAOTHER, # cancer diagnoses
    HYSTER53, PAPSMR53, #Cervical Cancer
    BSTST53, BSTSRE53, CLNTST53, CLNTRE53,SGMTST53,SGMTRE53 , # Colorectal Cancer
    MAMOGR53, # Breast Cancer
    PSA53 # Prostate Cancer:
)


```

for education half of the panel is following the year before thus need to be fixed
```{r}


fyc15$EDRECODE<-replace(fyc15$EDRECODE, fyc15$EDRECODE %in% 0:12, 1)
fyc15$EDRECODE<-replace(fyc15$EDRECODE, fyc15$EDRECODE %in% 13:14, 2)
fyc15$EDRECODE<-replace(fyc15$EDRECODE, fyc15$EDRECODE==15, 3)
fyc15$EDRECODE<-replace(fyc15$EDRECODE, fyc15$EDRECODE==16, 4)

fyc15<-rename(fyc15, EDUC=EDRECODE)

table(fyc15$EDUC)
# 1: No Dgree

# 2: GED or High school  2,3

# 3: Bachelors

# 4: Master's and doctorate


```

 
#2014

 There was no  HIDEG in this year data there is info about in documanation a subsittue variable used called
 EDRECODE read further in : 
https://meps.ahrq.gov/data_stats/download_data/pufs/h155/h155doc.shtml#2582FamilyOrigins

```{r}


fyc14  <- fyc2014 %>% select(
    DUPERSID,DUID,PANEL, PID,PSTATS31 , PSTATS42 , PSTATS53 , PERWT, VARSTR, VARPSU,SPOUID31, # IDS and weights
    AGE, SEX, RACETHX,HISPANX, MARRYX, FAMS1231,  EMPST31, POVCAT,POVLEV, DOBYY ,REGION, EDRECODE , # Demographic
    BORNUSA, YRSINUS, OTHLANG,LANGSPK, HWELLSPE, ADLANG42, # Immigration
    INSURC,MCDEV, MCREV, PRVEV, UNINS,INSCOV, PRIV ,#Insurance
    CABREAST, CACERVIX, CACOLON, CAPROSTA, CANCERDX, CAOTHER, # cancer diagnoses
    HYSTER53, PAPSMR53, #Cervical Cancer
    BSTST53, BSTSRE53, CLNTST53, CLNTRE53,SGMTST53,SGMTRE53 , # Colorectal Cancer
    MAMOGR53, # Breast Cancer
    PSA53 # Prostate Cancer:
)


```

education
```{r}


fyc14$EDRECODE<-replace(fyc14$EDRECODE, fyc14$EDRECODE %in% 0:12, 1)
fyc14$EDRECODE<-replace(fyc14$EDRECODE, fyc14$EDRECODE %in% 13:14, 2)
fyc14$EDRECODE<-replace(fyc14$EDRECODE, fyc14$EDRECODE==15, 3)
fyc14$EDRECODE<-replace(fyc14$EDRECODE, fyc14$EDRECODE==16, 4)

fyc14<-rename(fyc14, EDUC=EDRECODE)

table(fyc14$EDUC)
# 1: No Dgree

# 2: GED or High school  2,3

# 3: Bachelors

# 4: Master's and doctorate


```


#2013
 
 Neither EDUCYR nor HIDEG exist in this dataset instead there are: EDUYRDG, EDRECODE

```{r}


fyc13  <- fyc2013 %>% select(
    DUPERSID,DUID,PANEL, PID,PSTATS31 , PSTATS42 , PSTATS53 , PERWT, VARSTR, VARPSU,SPOUID31, # IDS and weights
    AGE, SEX, RACETHX,HISPANX, MARRYX, FAMS1231,  EMPST31, POVCAT,POVLEV, DOBYY ,REGION, EDRECODE, # Demographic
    BORNUSA,USBORN42, LIVEUS42,YRSINUS, OTHLANG,LANGSPK, HWELLSPE, ADLANG42, # Immigration
    INSURC,MCDEV, MCREV, PRVEV, UNINS,INSCOV, PRIV ,#Insurance
    CABREAST, CACERVIX, CACOLON, CAPROSTA, CANCERDX, CAOTHER, # cancer diagnoses
    HYSTER53, PAPSMR53, #Cervical Cancer
    BSTST53, BSTSRE53, CLNTST53, CLNTRE53,SGMTST53,SGMTRE53 , # Colorectal Cancer
    MAMOGR53, # Breast Cancer
    PSA53 # Prostate Cancer:
)

   

```
fixing the education 

```{r}

fyc13$EDRECODE<-replace(fyc13$EDRECODE, fyc13$EDRECODE %in% 0:12, 1)
fyc13$EDRECODE<-replace(fyc13$EDRECODE, fyc13$EDRECODE %in% 13:14, 2)
fyc13$EDRECODE<-replace(fyc13$EDRECODE, fyc13$EDRECODE==15, 3)
fyc13$EDRECODE<-replace(fyc13$EDRECODE, fyc13$EDRECODE==16, 4)

fyc13<-rename(fyc13, EDUC=EDRECODE)

# 1: No Dgree

# 2: GED or High school  2,3

# 3: Bachelors

# 4: Master's and doctorate


```

due to previous year change in US born some changes need to be done with this dataset:

```{r}
  # both US born variable mixed in one variable
        fyc13$BORNUSA<-ifelse(fyc13$PANEL==18,fyc13$BORNUSA,fyc13$USBORN42)
            # USBORN42 is extra and need to be dropped
    # same thing with years in US    
        fyc13$YRSINUS<-ifelse(fyc13$PANEL==18,fyc13$YRSINUS,fyc13$LIVEUS42)

    table(fyc13$PANEL,fyc13$BORNUSA)
    fyc13 = subset(fyc13, select = -c(USBORN42,LIVEUS42))

    
    
    
        table(fyc13$PANEL,fyc13$LIVEUS42)

            table(fyc13$PANEL,fyc13$YRSINUS)


```



#2012

Name change in US born and years lived in US. I didn't rename it yet because there might be differences.

```{r}


## 
fyc12  <- fyc2012 %>% select(
    DUPERSID,DUID,PANEL, PID,PSTATS31 , PSTATS42 , PSTATS53 , PERWT, VARSTR, VARPSU,SPOUID31, # IDS and weights
    AGE, SEX, RACETHX,HISPANX, MARRYX, FAMS1231,  EMPST31, POVCAT,POVLEV, DOBYY ,REGION,EDRECODE, # Demographic
    USBORN42, USLIVE42, ENGCMF42,ENGSPK42, LANGHM42, ADLANG42, # Immigration
    INSURC,MCDEV, MCREV, PRVEV, UNINS,INSCOV, PRIV ,#Insurance
    CABREAST, CACERVIX, CACOLON, CAPROSTA, CANCERDX, CAOTHER, # cancer diagnoses
    HYSTER53, PAPSMR53, #Cervical Cancer
    BSTST53, BSTSRE53, CLNTST53, CLNTRE53,SGMTST53,SGMTRE53 , # Colorectal Cancer
    MAMOGR53, # Breast Cancer
    PSA53 # Prostate Cancer:
)



## there should be a change in YLINU afterward USLIVE42 can be dropped



```

Further investigation showed that years in us was not only different name but it was also continuous while for after 2013 it was categorical data 
so I changed it as follow:

First, renamed the USBORN42 and  USLIVE42 to make it consistent with the rest
```{r}

  fyc12 <- fyc12 %>%
            rename(
                BORNUSA  = USBORN42,
                YRSINUS= USLIVE42 ,
                EDUC=EDRECODE
                ) 
```


then make YRSINUS categorical like other years

```{r}
        fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS %in% 1:4,2)  # 2 1 YR., LESS THAN 5 YRS.
        fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS %in% 5:9,3)  # 3 5 YRS., LESS THAN 10 YRS.
        fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS %in% 10:14,4)  # 4 10 YRS., LESS THAN 15 YRS.
        fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS >=15,5)  # 5 15 YEARS OR MORE
        fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS==0,1)  # 1 LESS THAN 1 YEAR

```

fix the education
```{r}

fyc12$EDUC<-replace(fyc12$EDUC, fyc12$EDUC %in% 0:12, 1)
fyc12$EDUC<-replace(fyc12$EDUC, fyc12$EDUC %in% 13:14, 2)
fyc12$EDUC<-replace(fyc12$EDUC, fyc12$EDUC==15, 3)
fyc12$EDUC<-replace(fyc12$EDUC, fyc12$EDUC==16, 4)

# 1: No Dgree

# 2: GED or High school  2,3

# 3: Bachelors

# 4: Master's and doctorate



```


#2011

In addition to the US-Born and everything there is a difference with RACE variable as well 

```{r}


fyc11  <- fyc2011 %>% select(
    DUPERSID,DUID,PANEL, PID,PSTATS31 , PSTATS42 , PSTATS53 , PERWT, VARSTR, VARPSU,SPOUID31, # IDS and weights
    AGE, SEX, RACEX,HISPANX, MARRYX, FAMS1231,  EMPST31, POVCAT,POVLEV, DOBYY ,REGION,EDRECODE, # Demographic
    USBORN42, USLIVE42, ENGCMF42,ENGSPK42, LANGHM42, ADLANG42, # Immigration
    INSURC,MCDEV, MCREV, PRVEV, UNINS,INSCOV, PRIV ,#Insurance
    CABREAST, CACERVIX, CACOLON, CAPROSTA, CANCERDX, CAOTHER, # cancer diagnoses
    HYSTER53, PAPSMR53, #Cervical Cancer
    BSTST53, BSTSRE53, CLNTST53, CLNTRE53,SGMTST53,SGMTRE53 , # Colorectal Cancer
    MAMOGR53, # Breast Cancer
    PSA53 # Prostate Cancer:
)
         


```

 start with the US borns: 

Same as 2012 first rename the USBORN42 and  USLIVE42 to make it consistent with the rest
```{r}
  fyc11<- fyc11 %>%
                rename(
                    BORNUSA= USBORN42,
                    YRSINUS= USLIVE42,
                    RACETHX=RACEX,
                    EDUC=EDRECODE) 
```

then make the USlive categorical 

```{r}

            fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS %in% 1:4,2)  # 2 1 YR., LESS THAN 5 YRS.
            fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS %in% 5:9,3)  # 3 5 YRS., LESS THAN 10 YRS.
            fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS %in% 10:14,4)  # 4 10 YRS., LESS THAN 15 YRS.
            fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS >=15,5) # 5 15 YEARS OR MORE
            fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS==0,1)  #  1 LESS THAN 1 YEAR

```

Now move on to the race variable, let's merge ethnicity into the race


```{r}
fyc11$RACETHX<-fyc11$RACETHX+1
fyc11$RACETHX<-replace(fyc11$RACETHX,fyc11$RACETHX %in% c(4, 6, 7),6)  # 3 5 YRS., LESS THAN 10 YRS.
fyc11$RACETHX<-replace(fyc11$RACETHX,fyc11$RACETHX==5,4)  # 3 5 YRS., LESS THAN 10 YRS.
fyc11$RACETHX<-replace(fyc11$RACETHX,fyc11$RACETHX==6,5)  # 3 5 YRS., LESS THAN 10 YRS.

fyc11$RACETHX<-ifelse(fyc11$HISPANX==1,1,fyc11$RACETHX)



```
fixing the education creating a new variable EDUC for all years

```{r}



fyc11$EDUC<-replace(fyc11$EDUC, fyc11$EDUC %in% 0:12, 1) # 1: No Dgree
fyc11$EDUC<-replace(fyc11$EDUC, fyc11$EDUC %in% 13:14, 2) # 2: GED or High school  2,3

fyc11$EDUC<-replace(fyc11$EDUC, fyc11$EDUC==15, 3) # 3: Bachelors
fyc11$EDUC<-replace(fyc11$EDUC, fyc11$EDUC==16, 4) # 4: Master's and doctorate


table(fyc11$PANEL, fyc11$EDUC)









```





