---
title: "R Notebook"
output: html_notebook
---

I started to find out what variables I need, an then I realized that the panel version of my research might be problematic
and I want to check that how could it be , so I'm looking at the info of panel 20 in year 15 and year 16

```{r}

yr15p20 <- fyc15 %>% 
  filter(PANEL == 20)


yr16p20 <- fyc16 %>% 
  filter(PANEL == 20)

subyr15p20 <- yr15p20[yr15p20$DUPERSID %in% yr16p20$DUPERSID, ]


```

I want to check how many unique ID i have in each dataset

```{r}

length(unique(yr15p20$DUPERSID))

length(unique(yr16p20$DUPERSID))
length(unique(subyr15p20$DUPERSID))


```
there might be some IDs in yr16p20 that are not present in yr15p20 I want to see what is happening

```{r}
missing_ids <- yr16p20[!yr16p20$DUPERSID %in% yr15p20$DUPERSID, ]
nrow(missing_ids)

```
now i need to delet this people from the sample 2016 so that I would have the exact same people in my panel
```{r}
subyr16p20 <- yr16p20[!yr16p20$DUPERSID %in% missing_ids$DUPERSID, ]


```


so this shows me that about 1200 person dropped out of the study in the next year
Now, I want to see if the report of screening is the same for individuals

```{r}
table(subyr15p20$CLNTST53)
table(subyr16p20$CLNTST53)


```

Now the  result is not close to what I expected U want to see if the people who said that I did colonscopy 

I'm going to see if the people how report that last time they colonscopy was 5 years ago the same as people who report as Never

```{r}


subyr15p20 <- subyr15p20 %>% 
  filter(CLNTST53 == 5)

subyr16p20 <- subyr16p20 %>% 
  filter(CLNTST53 == 5)

```

now check the difference between one person in one year and second year

```{r}


data1 <- yr15p20 %>% 
  filter(DUPERSID == 70389102)

data2 <- yr16p20 %>% 
  filter(DUPERSID == 70389102)

```

```{r}

table(data1$AGE31X)
table(data2$AGE31X)

table(data2$AGE53X)
table(data2$CLNTST53)
table(data1$CLNTST53)

```
In general what I figured out is that for cancer screening I can bring info from round 5 

## 1. Cleaning Data

### Load data
if I decided to the loop this code can help me
```{r}
### doing this in one for loop instead

# Create an empty list to store the results
fyc_list <- list()
# Loop through the years and read_MEPS for each year
for (year in 2011:2019) {
    # Read the data for the specific year
    fyc_data <- read_MEPS(year = year, type = "FYC")
    
    # Add a "year" variable to the dataset
    fyc_data$year <- year
    
    # Store the dataset in the list
    fyc_list[[paste0("fyc", year)]] <- fyc_data
}

# Combine all datasets into one big dataset

fyc <- do.call(rbind, fyc_list)
```


```{r}
d2019 <- read_MEPS(file = "h216")
p2019 <- read_MEPS(file = "h217")
d2019p <- read_MEPS(file = "h209")
pnl20<-read_MEPS(file = "h193")
```

another way of loading data set

```{r}
# FYC = Full year consolidated file (record = MEPS sample person)
fyc11 <- read_MEPS(year = 2011, type = "FYC") # Make this code a loop
fyc12 <- read_MEPS(year = 2012, type = "FYC")
fyc13 <- read_MEPS(year = 2013, type = "FYC")
fyc14 <- read_MEPS(year = 2014, type = "FYC")
fyc15 <- read_MEPS(year = 2015, type = "FYC")
fyc16 <- read_MEPS(year = 2016, type = "FYC")
fyc17 <- read_MEPS(year = 2017, type = "FYC")
fyc18 <- read_MEPS(year = 2018, type = "FYC")
fyc19 <- read_MEPS(year = 2019, type = "FYC")

```




### Choosing variables to subset  : 
April 22 11:04

Today I'm trying to prepare my data for the power analysis. I determined there are many variables that I need to determine later on about their inclusion  based on my models I'll includ them here for the later use.

```{r}
DDNWRK19, ADDAYA42, ADGENH42 , RTHLTH31, RTHLTH42,RTHLTH53, MNHLTH31, MNHLTH42,MNHLTH53, # Health
                             ADRTCR42,
                             ADSMOK42,  # risk
                             DLAYCA42, AFRDCA42, DLAYPM42,AFRDPM42, OBDEXP19,  # cost
                             HAVEUS42,TMTKUS42, PHNREG42, AFTHOU42, OBTOTV19, OBVTCH19, PRVSPK42, PROVTY42_M18 , YNOUSC42_M18,
                             ADILWW42, ADILCR42, ADRTWW42, # Aceess
                             OBDRV19, OBDTCH19 ,# healtj use
                             ADHECR42, ADLIST42, ADEZUN42, ADEZUN42, ADRESP42, ADPRTM42, # Perceived quality of health services 
                             ADINSA42 , ADINSB42 ,ADRISK42,ADOVER42, # distrust & risk
                                      #health behaviors
                             CHOLDX,  # Comorbidity
                             
                             REFRL19X, # for undoc
```

For cancer screening previously MEPS was gathering info for indiviual in round 3 and 5 basivly annually, but later after 2016 for each indiviual they asked it only once which make sense since some wouldn't change if they aren't current.I'll  be using their answer from last year for this analysis.

### Create a consistent variable for screening for all years

```{r}

ADBRST42, ADMMGR42 , # Breast Cancer
CABREAST
table(fyc18$ADBRST42)
table(fyc18$CANCERDX)
table(fyc18$AGE)

table(fyc18$PANEL)

```

let me first create a same variable for screening for years 2011 through 2016 then will move forward with creating similar variables for years with no information. Probably prior to do that I might need to fix the race variable for 2011 check if education need to be fixed or  US born requires any fixation.
starting 2013 the US born question changed place and therefore variable name changed in 2013 we have both USborn and BornUS  for each variable that needed to be fixed take a look and fix:

```{r}
table(fyc13$USBORN42)
table(fyc13$BORNUSA)

table(fyc13$PANEL, fyc13$USBORN42)
table(fyc13$PANEL, fyc13$BORNUSA)


fyc13$BORNUSA<-ifelse(fyc13$PANEL==18,fyc13$BORNUSA,fyc13$USBORN42)
fyc13$YRSINUS<-ifelse(fyc13$PANEL==18,fyc13$YRSINUS,fyc13$LIVEUS42)

table(fyc13$PANEL, fyc13$YRSINUS)
table(fyc13$PANEL, fyc13$LIVEUS42)

```
Another change in variable was the years lived in US, before 2013 this variable was a continuous variable named USLIVE42 however after 2013 it become a categorical variable YRSINUS. To be consistent I need to create a  categorical YRSINUS for years 2011 and 2012 and drop the USLIVE42 from these year data set here is the change:

Fixing the years in US

```{r}

table(fyc13$PANEL, fyc13$YRSINUS)
table(fyc14$PANEL, fyc14$YRSINUS)
table(fyc12$PANEL, fyc12$USLIVE42)

table(fyc12$USLIVE42)

fyc12$YRSINUS<-fyc12$USLIVE42

fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS %in% 1:4,2)  # 0: Single
fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS %in% 5:9,3)  # 0: Single
fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS %in% 10:14,4)  # 0: Single
fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS >=15,5)  # 0: Single
fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS==0,1)  # 0: Single

table(fyc11$YRSINUS)
table(fyc11$USLIVE42)
fyc11$YRSINUS<-fyc11$USLIVE42
fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS %in% 1:4,2)  # 0: Single
fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS %in% 5:9,3)  # 0: Single
fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS %in% 10:14,4)  # 0: Single
fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS >=15,5)
fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS==0,1)  # 0: Single



```
I want to find duplicated ID

```{r}
data<-rbind(fyc17,fyc18)


b<-data$DUPERSID[duplicated(data$PANEL)]
```
checking for loop
```{r}

for (year in 2015:2016) {
    
    dataset_name <- paste0("fyc", year) 

    assign(dataset_name, read_MEPS(year = year, type = "FYC"))  ## loading dataset
    
    a<-print(colnames(get(dataset_name)))

}
```
## checking the different between two dataset col names?

```{r}

d1<-colnames(fyc11)
d2<-colnames(fyc18)

d1<- columns_to_select

d2<- c(
    "DUPERSID", "DUID", "PANEL", "PID", "PSTATS31", "PSTATS42", "PSTATS53", "PERWT", "VARSTR", "VARPSU", "SPOUID31", # IDS and weights
    "AGE", "SEX", "RACETHX", "HISPANX", "MARRYX", "FAMS1231", "EMPST31", "POVCAT", "POVLEV", "DOBYY", "REGION", "EDRECODE", # Demographic
    "BORNUSA", "USBORN42", "LIVEUS42", "YRSINUS", "ENGSPK42", "HWELLSPE", # Immigration
    "INSURC", "MCDEV", "MCREV", "PRVEV", "UNINS", "INSCOV", "PRIV", # Insurance
    "CABREAST", "CACERVIX", "CACOLON", "CAPROSTA", "CANCERDX", "CAOTHER", # Cancer diagnoses
    "HIBPDX", "DIABDX", "ASTHDX", "CHDDX", "ARTHDX", "CANCERDX", # Comorbidities
    # "HYSTER53", "PAPSMR53", # Cervical Cancer
    "CLNTST53", "SGMTST53",  # Colorectal Cancer (dropped CLNTRE53, SGMTRE53)
    "MAMOGR53",  # Breast Cancer
    # "PSA53", # Prostate Cancer
    "ADRISK42", "ADOVER42", "HAVEUS42", "RTHLTH42", "ADSMOK42", "PHYEXE53" # Health and health behavior
)

setdiff(d1,d2)
setdiff(d2,d1)




setdiff(d1,d2)
[1] "ADRISK42" "ADOVER42" "ADSMOK42"
> 
> setdiff(d2,d1)
[1] "HWELLSPE" "CLNTST53" "SGMTST53" "MAMOGR53" "YEAR" 

```

```{r}

fyc19<-fyc2019
fyc19$CLNTST53<-99999

######################## from 2020 #############
matching_indices <- match(fyc19$DUPERSID, fyc2020$DUPERSID)

# Create a logical vector indicating non-missing matching indices
valid_matching_indices <- !is.na(matching_indices)

# Get the corresponding PANEL values for the matched indices
panel_values <- fyc2020$PANEL[matching_indices[valid_matching_indices]]

# Find the indices where PANEL is 24
panel_24_indices <- panel_values == 24

# Use these indices to subset the replacement values correctly
replacement <- fyc2020$CLNTST53[matching_indices[valid_matching_indices][panel_24_indices]]

# Now replace the values in fyc19
fyc19$CLNTST53 <- replace(fyc19$CLNTST53, valid_matching_indices & fyc19$PANEL == 24, replacement)

######################## from 2018 #############


matching_indices <- match(fyc19$DUPERSID, fyc18$DUPERSID)

# Create a logical vector indicating non-missing matching indices
valid_matching_indices <- !is.na(matching_indices)

# Get the corresponding PANEL values for the matched indices
panel_values <- fyc18$PANEL[matching_indices[valid_matching_indices]]

# Find the indices where PANEL is 24
panel_23_indices <- panel_values == 23

# Use these indices to subset the replacement values correctly
replacement <- fyc18$CLNTST53[matching_indices[valid_matching_indices][panel_23_indices]]

# Now replace the values in fyc19
fyc19$CLNTST53 <- replace(fyc19$CLNTST53, valid_matching_indices & fyc19$PANEL == 23, replacement)



table(fyc19$CLNTST53)

```

create a function that generate table for list of variables

```{r}

variables_to_check <- c('FAMS1231', 'EMPST31', 'BORNUSA', 'YRSINUS', 'ENG', 'CLNTST53', 'SGMTST53', 'MAMOGR53')

FAMS1231, EMPST31, BORNUSA, YRSINUS, ENG, CLNTST53, SGMTST53, MAMOGR53, HIBPDX, DIABDX, ASTHDX, CHDDX, ARTHDX, CANCERDX
 HAVEUS42, RTHLTH42, PHYEXE53, ADRISK42, ADOVER42, ADSMOK42
 

generate_table <- function(dataset, var_name) {
  table(dataset[[var_name]])
}

tables_list <- lapply(variables_to_check, function(var) generate_table(data, var))
names(tables_list) <- variables_to_check

for (var in names(tables_list)) {
  cat("\nTable for", var, ":\n")
  print(tables_list[[var]])
}

```
check how are the people how have no clonscopy 

```{r}

check <- data %>%
  filter(ADSMOK42 == -1) 

check1 <- fyc16 %>%
  filter(DUPERSID==2110806109) 

data$CANCERDX

bindraw<- fyc2018 %>% filter(AGE >49 & AGE<65 )

table(check$YEARcheck$CHDDX)
table(check$AGE)
table(check$YEAR)

table(check$SAQELIG)
table(check$SAQELIG,check$SGMTST53)

table(check$MAMOGR53, check$SEX)

```

## changing 2 to 0 for dummy variables 

I want to create a function to change 2 to 0 for my variables, Two way I wrote the function using the 

Base 

```{r}

dummy <- function(df, col) {
  df[[col]] = as.numeric ( df[[col]])
  df[[col]][df[[col]] == 2] <- 0
  return(df)
}

# Modify data using dummy function for SEX column only

# data <- dummy(data, "SEX")

# table(data$SEX)

```



using dplyr 


```{r}

library(dplyr)

dummy <- function(df, col) {
  #col <- enquo(col)  # Convert the column name to a quosure
  col <- sym(col)
  df <- df %>%
    mutate(
      {{ col }} := as.numeric ( {{ col }}),
      {{ col }} := case_when(
      {{ col }} == 2 ~ 0,
      TRUE ~ {{ col }}
    ))
  
  return(df)
}



```

check my function

```{r}

for (var in c("SEX", "UNINS")) {
  data <- dummy(data, var)
}

```

 look at the table of all variables
 
```{r}


for (var in Allvariables){
 data %>% 
    count(var)
  
}

```
 

