\documentclass[../main.tex]{subfiles}

\begin{document}





\end{document}

table captions: 
Event-study plots with controls
Effect of Medicaid expansion, by Nativity

The figure plots estimates of the effect of the closing on enrollments in county with closing in the years before and after the closing. Specifically, the figure plots estimates of ˆ δℓ coefficients from equation (2.4), which is a interaction weighted estimator of the log of enrollments on county fixed effects, state specific year fixed effects, control covariates and closing event year indicators. The dependent variable is the log number of enrolled students. The marker is depicted along with the 95 percent confidence intervals. Source: NCES IPEDS survey data on institutions and Unemployment data by BEA.

te: The figure plots estimates of ˆ δℓ coefficients from equation (2.4) for subgroups. The dependent variable is the log number of enrolled students by subgroups. The marker is depicted along with the 95 percent confidence intervals. Source: NCES IPEDS survey data on institutions.


Table presents the estimate of the effect of the closing on enrollments in counties with closing in the years before and after the closing. Specifically, the table reports the estimates of ˆ δℓ coefficients from equation (2.4), which is estimated with county fixed effects, state specific year fixed effects, control covariates and closing event year indicators. The dependent variable is the log number of enrolled students. In panel A, the estimation sample includes all never-treated counties and counties with one-time for-profit closure. In panel B, the sample includes all counties with less than 15 institutions. In panel C, the sample is the same as our baseline sample but closing indicators are defined by a for-profit closing event in state-level. Standard errors are in parentheses and clustered in county level for panel A and B, and in state for panel C. Source: NCES IPEDS survey data on institutions from 2010 to 2019