---
title: "01-data-cleaning"
output: html_notebook
---


# About this code 

This is first attempt for data cleaning for paper 2

```{r,eval=FALSE}

ADRISK42, ADOVER42, HAVEUS42, RTHLTH42, ADSMOK42, PHYEXE53


ADRISK42 – More likely to take risks than the average person

ADOVER42 – Can overcome illness without help from a medically trained person

HAVEUS42- Have usual source of care 

RTHLTH42- PERCEIVED HEALTH STATUS

ADSMOK42- Currently smoke

PHYEXE53- physical activity at least five times a week Age>17; 

HIBPDX,DIABDX,ASTHDX,CHDDX, ARTHDX,CANCERDX


```


```{r, eval=FALSE}
# Install the required packages

install.packages("MEPS")
install.packages("survey")
install.packages("tidyverse")
install.packages("haven")
install.packages("labelled")
install.packages("broom")
install.packages("foreign")  
install.packages("devtools")
devtools::install_github("e-mitchell/meps_r_pkg/MEPS")
install.packages("fixest")

```


## 1. Loading required libraries

```{r}
library(foreign)
library(devtools)
library(survey) # handeling survey package
library(tidyverse)
library(haven)
library(labelled)
library(broom)
library(MEPS)  # readin MEPS data
library(readxl) # reading 
library(fixest)

```

```{r}

wd <- list()
# commonly used paths in my working directory
wd$data   <- "/home/shadi/Projects/GitHub/ACA-immigrant-preventive/data/"
wd$output <- "/home/shadi/Projects/GitHub/ACA-immigrant-preventive/output/"
wd$texts <- "/home/shadi/Projects/GitHub/ACA-immigrant-preventive/text/"
wd$codes <- "/home/shadi/Projects/GitHub/ACA-immigrant-preventive/code/R/"


```

## 2. Set survey option for lonely PSUs
```{r}
options(survey.lonely.psu="adjust")

```



## 3. Loading Dataset  

Variable for initial subseting 


```{r}

columns_to_select <- c(
    'YEAR', "DUPERSID", "DUID", "PANEL", "PID", "PSTATS31", "PSTATS42", "PSTATS53", "PERWT", 
    "VARSTR", "VARPSU", "SPOUID31", # IDS and weights
    "AGE", "SEX", "RACETHX", "HISPANX", "MARRYX", "FAMS1231", "EMPST31", "POVCAT", 
    "POVLEV", "DOBYY", "REGION", "EDUC", # Demographic
    "BORNUSA", "YRSINUS", "HWELLSPE", 'HWELLSPK','ENGSPK42',"ENG", 'OTHLANG', "OTHLGSPK","LANGHM42", "USBORN42", "LIVEUS42", # Immigration
    "INSURC", "MCDEV", "MCREV", "PRVEV", "UNINS", # Insurance
    "CABREAST","CACOLON", # "CACERVIX",  "CAPROSTA", "CANCERDX", "CAOTHER", # Cancer diagnoses
    "CLNTST53", "SGMTST53", # Colorectal Cancer
    "MAMOGR53", # Breast Cancer
    "HIBPDX", "DIABDX", "ASTHDX", "CHDDX", "ARTHDX", "CANCERDX", # Comorbidities
    "HAVEUS42", "RTHLTH42", "PHYEXE53", # Health and health behavior 
    "VPCS42", "VMCS42","PCS42","MCS42",
    "ADRISK42", "ADOVER42","ADSMOK42"
)

```


Load datasets for all years, rename year-specific variables and subset the needed variables for cleaning prior to combining  

```{r}


for (year in 2011:2019) {
    
    dataset_name <- paste0("fyc", year) 

    assign(dataset_name, read_MEPS(year = year, type = "FYC"))  ## loading dataset
    

            # renaming variable in dataset 
    base_renames <- get(dataset_name) %>%
           rename(
                   PERWT  = paste0("PERWT", substr(year, 3, 4), "F"), # substr(), the function extracts a substring from the year, starting from the third 
                                                                        #character position and ending at the fourth character position
                   AGE = paste0("AGE", substr(year, 3, 4), "X"),
                   MARRYX = paste0("MARRY", substr(year, 3, 4), "X"),
                   POVCAT = paste0("POVCAT", substr(year, 3, 4)),
                   POVLEV = paste0("POVLEV", substr(year, 3, 4) ),
                   REGION = paste0("REGION", substr(year, 3, 4)),
                   INSURC = paste0("INSURC", substr(year, 3, 4)), #Insurance
                   MCDEV = paste0("MCDEV", substr(year, 3, 4)), 
                   MCREV = paste0("MCREV", substr(year, 3, 4)), 
                   PRVEV = paste0("PRVEV", substr(year, 3, 4)) , 
                   UNINS = paste0("UNINS", substr(year, 3, 4)) ,
                   INSCOV = paste0("INSCOV", substr(year, 3, 4)) ,
                   PRIV = paste0("PRIV", substr(year, 3, 4))  
               )
    # Additional renaming for specific years
    
       if (year == 2019) {          # Add other renames specific to 2019
              base_renames <- base_renames %>%
                  rename(
                      DIABDX = DIABDX_M18,
                      EDUC = HIDEG ,
                      PCS42 = VPCS42,
                      MCS42 = VMCS42

                  )
       } else if (year == 2018) {       #renames specific to 2019
              base_renames <- base_renames %>%
                   rename(
                      DIABDX   = DIABDX_M18,
                      EDUC = HIDEG,
                      CLNTST53 = ADCLNS42,
                      SGMTST53 = ADSGMD42,
                      MAMOGR53 = ADMMGR42,
                      PCS42 = VPCS42,
                      MCS42 = VMCS42

                  )  
       } else if (year == 2017) {  # Rename specific to 2017
        base_renames <- base_renames %>%
            rename(
                EDUC = HIDEG,
                PCS42 = VPCS42, # Rename specific to 2017
                MCS42 = VMCS42
            )
    } else if (year %in% c(2016)) {
        base_renames <- base_renames %>%
            rename(
                EDUC = HIDEG
            )

      }  else if (year %in% c(2013,2014,2015)) {
              base_renames <- base_renames %>%
                   rename(
                     EDUC=EDRECODE
                            )
      } else if (year == 2012) {       #renames specific to 2019
              base_renames <- base_renames %>%
                   rename(  # rename the USBORN42 and  USLIVE42 , language to make it consistent with the rest of dataset
                      BORNUSA = USBORN42, 
                      YRSINUS = USLIVE42,
                      EDUC = EDRECODE,
                      ENG = ENGSPK42
                  )  
     } else if (year == 2011) {       #renames specific to 2019
              base_renames <- base_renames %>%
                   rename(
                      BORNUSA = USBORN42,
                      YRSINUS = USLIVE42,
                      EDUC = EDRECODE,
                      ENG = ENGSPK42,
                      RACETHX = RACEX,
    
                  )  
     }

# apply the renaming     
assign(dataset_name,base_renames) 
  
assign(dataset_name, 
           get(dataset_name) %>%
               mutate(
                   YEAR = year
               )
      )

# select subset of variable (column) of datasets

selected_columns <- intersect(columns_to_select, colnames(get(dataset_name)))

new_data_selected <- get(dataset_name) %>%
        select(selected_columns)
assign(paste0("fyc", substring(year, 3,4)), new_data_selected)

}
   
```

years 2011 to 2019 are my timeline span, however I need year 2020 in order to create screening variable for one of the panel in 2019.

```{r}

fyc2020<-read_MEPS(year = 2020, type = "FYC")
fyc2020$YEAR<-2020
fyc2020<-fyc2020 %>% rename(
    CLNTST53 = ADCLNS42,
    SGMTST53 = ADSGMD42,
    MAMOGR53 = ADMMGR42
)
fyc2020$CLNTST53<-replace(fyc2020$CLNTST53, fyc2020$CLNTST53 %in% c(2,3),0) # 0 : No colonscopy within 10 years
fyc2020$SGMTST53<-replace(fyc2020$SGMTST53, fyc2020$SGMTST53 %in% c(2,3),0) # 0 : No semidoscopy within 5 years
fyc2020$MAMOGR53<-replace(fyc2020$MAMOGR53, fyc2020$MAMOGR53 %in% c(2,3),0) # 0 : No colonscopy within 2 years

# Create labeled factors for each variable

  fyc2020$CLNTST53 <- labelled( fyc2020$CLNTST53,
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "N/A" = -15) ,
                              label = "Had colonoscopy within 10 years"
                              )

    fyc2020$SGMTST53 <- labelled( fyc2020$SGMTST53,
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "N/A" = -15) ,
                              label = "Had semidoscopy within 5 years"
                              )

                      
    fyc2020$MAMOGR53 = labelled((fyc2020$MAMOGR53), 
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "N/A" = -15) ,
                              label = "Had mammogram within 2 years"
                              )






```

## 4. Cleaning variables


### Cancer Screening

for 2018 

```{r}
fyc18 <- fyc18 %>%
  mutate(
    # Replacing values
    CLNTST53 = replace(CLNTST53, CLNTST53 %in% c(2, 3), 0),
    SGMTST53 = replace(SGMTST53, SGMTST53 %in% c(2, 3), 0),
    MAMOGR53 = replace(MAMOGR53, MAMOGR53 %in% c(2, 3), 0)
  )


    # Assigning labels using haven::labelled
  fyc18$CLNTST53 <- labelled( fyc18$CLNTST53,
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "N/A" = -15) ,
                              label = "Had colonoscopy within 10 years"
                              )

    fyc18$SGMTST53 <- labelled( fyc18$SGMTST53,
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "N/A" = -15) ,
                              label = "Had semidoscopy within 5 years"
                              )

                      
    fyc18$MAMOGR53 = labelled((fyc18$MAMOGR53), 
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "N/A" = -15) ,
                              label = "Had mammogram within 2 years"
                              )

                      






```

for years between 2011-16


```{r}

datasets <- list(fyc11, fyc12, fyc13, fyc14, fyc15, fyc16)

for (i in seq_along(datasets)) {
  data <- datasets[[i]]
  
  # Colonscopy - CLNTST53
  data$CLNTST53 <- ifelse(data$AGE < 50, -1, data$CLNTST53)
  data$CLNTST53 <- ifelse(data$CACOLON == 1, -1, data$CLNTST53)
  data$CLNTST53 <- replace(data$CLNTST53, data$CLNTST53 %in% 1:5, 1)  # YES Colonscopy within 10 years
  data$CLNTST53 <- replace(data$CLNTST53, data$CLNTST53 %in% c(6, 7), 0) # NO
  data$CLNTST53 <- replace(data$CLNTST53, data$CLNTST53 %in% c(-9, -8,-7), -15) # N/A

  
    ### Add labels 
    data$CLNTST53 <- labelled( data$CLNTST53,
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "N/A" = -15) ,
                              label = "Had colonoscopy within 10 years"
                              )

  
  # Semidoscoy - SGMTST53
  data$SGMTST53 <- ifelse(data$AGE < 50, -1, data$SGMTST53)
  data$SGMTST53 <- ifelse(data$CACOLON == 1, -1, data$SGMTST53)
  data$SGMTST53 <- replace(data$SGMTST53, data$SGMTST53 %in% 1:4, 1)  # YES semidoscopy within 5 years
  data$SGMTST53 <- replace(data$SGMTST53, data$SGMTST53 %in% c(5, 6, 7), 0) # NO
  data$SGMTST53 <- replace(data$SGMTST53, data$SGMTST53 %in% c(-9, -8,-7), -15) # N/A

   ## add labels
      data$SGMTST53 <- labelled( data$SGMTST53,
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "N/A" = -15) ,
                              label = "Had semidoscopy within 5 years"
                              )

  
  # Mammogram - MAMOGR53
  data$MAMOGR53 <- ifelse(data$AGE < 50, -1, data$MAMOGR53)
  data$MAMOGR53 <- ifelse(data$CABREAST == 1, -1, data$MAMOGR53)
  data$MAMOGR53 <- replace(data$MAMOGR53, data$MAMOGR53 %in% 1:2, 1)  # YES Mammogram within 10 years
  data$MAMOGR53 <- replace(data$MAMOGR53, data$MAMOGR53 %in% 3:6, 0)  # NO
  data$MAMOGR53 <- replace(data$MAMOGR53, data$MAMOGR53 %in% c(-9, -8,-7), -15) # N/A
  
   ### Add labels
        data$MAMOGR53 = labelled((data$MAMOGR53), 
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "N/A" = -15) ,
                              label = "Had mammogram within 2 years"
                              )

  

  datasets[[i]] <- data
}

# Reassign the modified datasets back to the original variables
fyc11 <- datasets[[1]]
fyc12 <- datasets[[2]]
fyc13 <- datasets[[3]]
fyc14 <- datasets[[4]]
fyc15 <- datasets[[5]]
fyc16 <- datasets[[6]]

```

Imputing for year 2017 and 2019

```{r}
# Initialize columns with 99999
cols <- c("CLNTST53", "SGMTST53", "MAMOGR53")
fyc19[cols] <- -15
fyc17[cols] <- -15

# Update fyc19 columns from fyc2020
matching_indices <- match(fyc19$DUPERSID, fyc2020$DUPERSID)
valid_indices <- !is.na(matching_indices) & fyc19$PANEL == 24
fyc19[valid_indices, cols] <- fyc2020[matching_indices[valid_indices], cols]

# Update fyc19 columns from fyc18
matching_indices <- match(fyc19$DUPERSID, fyc18$DUPERSID)
valid_indices <- !is.na(matching_indices)
fyc19[valid_indices, cols] <- fyc18[matching_indices[valid_indices], cols]

# Fix naming difference for fyc17
fyc17$b <- paste0(as.character(fyc17$PANEL), fyc17$DUPERSID)

# Update fyc17 columns from fyc18
matching_indices <- match(fyc17$b, fyc18$DUPERSID)
valid_indices <- !is.na(matching_indices)
fyc17[valid_indices, cols] <- fyc18[matching_indices[valid_indices], cols]

# Update fyc17 columns from fyc16
matching_indices <- match(fyc17$DUPERSID, fyc16$DUPERSID)
valid_indices <- !is.na(matching_indices)
fyc17[valid_indices, cols] <- fyc16[matching_indices[valid_indices], cols]

# Remove the temporary column 'b'
fyc17 <- fyc17[ , !names(fyc17) %in% "b"]


```

### Education

for years 2016-2019
I need to make education consistent for all years

```{r}

datasets <- list(fyc16,fyc17,fyc18,fyc19)

for (i in seq_along(datasets)) {
  data <- datasets[[i]]
  
 data$EDUC<-replace(data$EDUC, data$EDUC %in% c(2,3,7),2)# 2: GED or High school  2,3
 data$EDUC<-replace(data$EDUC, data$EDUC == 4,3) # 3: Bachelors
 data$EDUC<-replace(data$EDUC, data$EDUC %in% c(5,6),4) # 4: Master's and doctorate
 data$EDUC<-replace(data$EDUC, data$EDUC %in% c(-7,-8,-9),-15) # NA
 
 
    ## add labels
      data$EDUC <- labelled( data$EDUC,
                              labels = c("No Degree" = 1 ,
                                          "GED/High School" = 2, 
                                         "Bachelors" = 3 , 
                                         "Master/Doctorate" = 4,
                                         INAPPLICABLE = 8 , 
                                         "N/A" = -15) ,
                              label = "Education Level"
                              )


  datasets[[i]] <- data
}







# Reassign the modified datasets back to the original variables
fyc16 <- datasets[[1]]
fyc17 <- datasets[[2]]
fyc18 <- datasets[[3]]
fyc19 <- datasets[[4]]


```

There was no  HIDEG in 2014 data there is info about in documanation a subsittue variable used called
 EDRECODE read further in : 
https://meps.ahrq.gov/data_stats/download_data/pufs/h155/h155doc.shtml#2582FamilyOrigins

for years 2011-2015

```{r}
datasets <- list(fyc11, fyc12, fyc13, fyc14, fyc15)

for (i in seq_along(datasets)) {
  data <- datasets[[i]]
  
  data$EDUC<-replace(data$EDUC, data$EDUC %in% 0:12, 1) # 1: No Dgree
  data$EDUC<-replace(data$EDUC, data$EDUC %in% 13:14, 2) # 2: GED or High school  2,3
  data$EDUC<-replace(data$EDUC, data$EDUC==15, 3) # 3: Bachelors
  data$EDUC<-replace(data$EDUC, data$EDUC==16, 4) # 4: Master's and doctorate
 data$EDUC<-replace(data$EDUC, data$EDUC %in% c(-7,-8,-9),-15) # NA

  
     ## add labels
      data$EDUC <- labelled( data$EDUC,
                              labels = c("No Degree" = 1 ,
                                          "GED/High School" = 2, 
                                         "Bachelors" = 3 , 
                                         "Master/Doctorate" = 4,
                                         INAPPLICABLE = -1 , 
                                         "N/A" = -15) ,
                              label = "Education Level"
                              )

  
  
  datasets[[i]] <- data
}

# Reassign the modified datasets back to the original variables
fyc11 <- datasets[[1]]
fyc12 <- datasets[[2]]
fyc13 <- datasets[[3]]
fyc14 <- datasets[[4]]
fyc15 <- datasets[[5]]


```


### Language

for year 2019
```{r}

fyc19$ENG<- replace(fyc19$HWELLSPK,fyc19$HWELLSPK %in% c(1,2), 0) # comfortable speaking English
fyc19$ENG<- replace(fyc19$ENG,fyc19$ENG %in% c(3,4), 1 ) #  not comfortable speaking English
fyc19$ENG<- ifelse(fyc19$OTHLGSPK==2, 0, fyc19$ENG)  #  # imputing -1 values to comfortable if person didn't speak any other language rather than Eng at hom
fyc19$ENG<- replace(fyc19$ENG,fyc19$ENG %in% c(-8,-9,-15), -15 ) # NA 

## add lable 

        fyc19$ENG = labelled((fyc19$ENG), 
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "Under 5 years old" = 5,
                                         "N/A" = -15) ,
                              label = "Not Comfortable speaking english"
                              )


fyc19 <- fyc19 %>% select(-HWELLSPK, -OTHLGSPK )

```

for year 2018 

this year again has different questionnaire for panels
change it to binary that answer if the person is not comfortable speaking English
```{r}
fyc18$ENG<- ifelse(fyc18$PANEL== 22, fyc18$HWELLSPE,fyc18$HWELLSPK)
fyc18$ENG<- replace(fyc18$ENG,fyc18$ENG %in% c(1,2), 0 ) # No, comfortable speaking English
fyc18$ENG<- replace(fyc18$ENG,fyc18$ENG %in% c(3,4), 1 ) # Yes, Not comfortable speaking English
fyc18$ENG<- replace(fyc18$ENG,fyc18$ENG %in% c(-8,-9,-15), -15 ) # NA 

# imputing -1 values to comfortable if person didn't speak any other language rather than eng at home

fyc18$ENG<- ifelse(fyc18$OTHLANG==2|fyc18$OTHLGSPK==2, 0, fyc18$ENG) 
fyc18$ENG<- ifelse(fyc18$BORNUSA==1 &  fyc18$ENG==-1 , 0, fyc18$ENG) 


# # imputing -1 values to comfortable if person didn't speak any other language rather than eng at home

# add labels 
        fyc18$ENG = labelled((fyc18$ENG), 
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "Under 5 years old" = 5,
                                         "N/A" = -15) ,
                              label = "Not Comfortable speaking english"
                              )


fyc18 <- fyc18 %>% select(-HWELLSPE, -HWELLSPK , -OTHLGSPK, -OTHLANG)

```

for year 2014-2017

change it to binary that answer if the person is not comfortable speaking English
```{r}

datasets <- list(fyc14, fyc15, fyc16, fyc17)

for (i in seq_along(datasets)) {
  data <- datasets[[i]]
  
  data$ENG<- replace(data$HWELLSPE,data$HWELLSPE %in% c(1,2), 0 ) # No, comfortable speaking English
  data$ENG<- replace(data$ENG,data$ENG %in% c(3,4), 1 ) # Yes, not comfortable
  data$ENG<- replace(data$ENG,data$ENG %in% c(-8,-9,-7,-15), -15 ) # NA 

# imputing -1 values to comfortable if person didn't speak any other language rather than Eng at home
  data$ENG<- ifelse(data$OTHLANG==2, 0, data$ENG)
  data$ENG<- ifelse(data$BORNUSA==1 & data$ENG==-1, 0, data$ENG) 

  # add labels 
        data$ENG = labelled((data$ENG), 
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "Under 5 years old" = 5,
                                         "N/A" = -15) ,
                              label = "Not Comfortable speaking english"
                              )

  
  

  data <-data  %>%  select(-HWELLSPE, - OTHLANG)

  datasets[[i]] <- data
}

# Reassign the modified datasets back to the original variables
fyc14 <- datasets[[1]]
fyc15 <- datasets[[2]]
fyc16 <- datasets[[3]]
fyc17 <- datasets[[4]]


```

for year 2013
change in questionare this year half dataset half different question make them the same

```{r}

fyc13$HWELLSPE<- replace(fyc13$HWELLSPE,fyc13$HWELLSPE %in% c(1,2), 2 ) # comfortable speaking english
fyc13$HWELLSPE<- replace(fyc13$HWELLSPE,fyc13$HWELLSPE %in% c(3,4), 1) # not comfortable
fyc13$HWELLSPE<- replace(fyc13$HWELLSPE,fyc13$HWELLSPE %in% c(-8,-9,-7,-15), -15 ) # NA 

fyc13$ENG<- ifelse(fyc13$PANEL== 18, fyc13$HWELLSPE,fyc13$ENGSPK42)
# imputing -1 values to comfortable if person didn't speak any other language rather than eng at home
fyc13$ENG<- replace(fyc13$ENG, fyc13$ENG==2, 0)

fyc13$ENG<- ifelse(fyc13$OTHLANG==2, 0, fyc13$ENG) 
fyc13$ENG<- ifelse(fyc13$OTHLANG %in% c(-8,-9,-7,-15), -15, fyc13$ENG) 

fyc13$ENG<- ifelse(fyc13$ENG==-1 & fyc13$BORNUSA==1, 0, fyc13$ENG) 
fyc13$ENG<- ifelse(fyc13$ENG==-1 & fyc13$USBORN42==1, 0, fyc13$ENG)

fyc13$ENG<- ifelse(fyc13$ENG==-1 & fyc13$USBORN42==1, 0, fyc13$ENG)

  # add labels 
        fyc13$ENG = labelled((fyc13$ENG), 
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "Under 5 years old" = 5,
                                         "N/A" = -15) ,
                              label = "Not Comfortable speaking english"
                              )





fyc13 <-fyc13  %>%  select(-HWELLSPE, -ENGSPK42, -OTHLANG, -LANGHM42)

```

for year 2011 & 2012

```{r}


fyc12$ENG<- replace(fyc12$ENG, fyc12$ENG==2, 0)
fyc12$ENG<- replace(fyc12$ENG, fyc12$ENG %in% c(-8,-9,-7,-15), -15)


fyc12$ENG<- ifelse(fyc12$ENG==-1 & fyc12$LANGHM42== 1, 0, fyc12$ENG) 
fyc12$ENG<- ifelse(fyc12$ENG==-1 & fyc12$BORNUSA==1, 0, fyc12$ENG) 

  # add labels 
        fyc12$ENG = labelled((fyc12$ENG), 
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "N/A" = -15) ,
                              label = "Not Comfortable speaking english"
                              )


fyc12 <-fyc12  %>%  select(-LANGHM42)


fyc11$ENG<- replace(fyc11$ENG, fyc11$ENG==2, 0)
fyc11$ENG<- replace(fyc11$ENG, fyc11$ENG %in% c(-8,-9,-7,-15), -15)

fyc11$ENG<- ifelse(fyc11$ENG==-1 & fyc11$LANGHM42== 1, 0, fyc11$ENG) 
fyc11$ENG<- ifelse(fyc11$ENG==-1 & fyc11$BORNUSA==1, 0, fyc11$ENG) 

  # add labels 
        fyc11$ENG = labelled((fyc11$ENG), 
                              labels = c(NO = 0, 
                                         YES = 1 , 
                                         INAPPLICABLE = -1,
                                         "N/A" = -15) ,
                              label = "Not Comfortable speaking english"
                              )



fyc11 <-fyc11  %>%  select(-LANGHM42)


```




### US-Born

for 2013

due to previous year change in US born some changes need to be done with this dataset:

```{r}
  # both US born variable mixed in one variable
        fyc13$BORNUSA<-ifelse(fyc13$PANEL==18,fyc13$BORNUSA,fyc13$USBORN42)
            # USBORN42 is extra and need to be dropped
    # same thing with years in US    
        fyc13$YRSINUS<-ifelse(fyc13$PANEL==18,fyc13$YRSINUS,fyc13$LIVEUS42)

    table(fyc13$PANEL,fyc13$BORNUSA)
    
    fyc13 = subset(fyc13, select = -c(USBORN42,LIVEUS42))



```
for all years

```{r}

fyc_names<- paste0("fyc", 11:19)
fycsets <- mget(fyc_names)
  
for (i in seq_along(fycsets)) {
 data <-fycsets[[i]]
 data <- data %>% 
    mutate (
        BORNUSA = replace(BORNUSA,BORNUSA %in% c(-7,-8,-9,-15), -15 ), # Missing
        BORNUSA = replace(BORNUSA,BORNUSA %in% c(1,-1), 0 ), # No, US Born 
        BORNUSA = replace(BORNUSA,BORNUSA %in% c(2), 1), # Yes, Foreign Born
        BORNUSA = labelled(
            BORNUSA, 
            labels = c(
                      "US-Born" = 0, 
                      "Foregin-Born" = 1 , 
                      INAPPLICABLE = -1,
                      "N/A" = -15
                    ) ,
            label = "Nativity Status"
            )
        )%>%
            rename(FORGNBORN = BORNUSA)

  assign(fyc_names[i], data)

}

```


### Years in US 

for year 2012

```{r}
        fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS %in% 1:4,2)  # 2 1 YR., LESS THAN 5 YRS.
        fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS %in% 5:9,3)  # 3 5 YRS., LESS THAN 10 YRS.
        fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS %in% 10:14,4)  # 4 10 YRS., LESS THAN 15 YRS.
        fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS >=15,5)  # 5 15 YEARS OR MORE
        fyc12$YRSINUS<-replace(fyc12$YRSINUS,fyc12$YRSINUS==0,1)  # 1 LESS THAN 1 YEAR

```

for year 2011

```{r}

            fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS %in% 1:4,2)  # 2 1 YR., LESS THAN 5 YRS.
            fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS %in% 5:9,3)  # 3 5 YRS., LESS THAN 10 YRS.
            fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS %in% 10:14,4)  # 4 10 YRS., LESS THAN 15 YRS.
            fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS >=15,5) # 5 15 YEARS OR MORE
            fyc11$YRSINUS<-replace(fyc11$YRSINUS,fyc11$YRSINUS==0,1)  #  1 LESS THAN 1 YEAR

```


```{r}

fyc_names<- paste0("fyc", 11:19)
fycsets <- mget(fyc_names)
  
for (i in seq_along(fycsets)) {
 data <-fycsets[[i]]
 data <- data %>% 
    mutate (
        YRSINUS = ifelse(YRSINUS ==-1, 0, YRSINUS), # 0 US-born 
        YRSINUS = replace(YRSINUS,YRSINUS %in% c(-7,-8,-9,-15), -15 ), # Missing
        YRSINUS = replace(YRSINUS,YRSINUS %in% c(1,2), 1 ), # Less than 5 Years
        YRSINUS = replace(YRSINUS,YRSINUS %in% c(3,4), 2 ), # Between 5-15 years
        YRSINUS = replace(YRSINUS,YRSINUS %in% c(5), 3 ), # More than 15 years
        YRSINUS = labelled(
            YRSINUS, 
            labels = c(
                      "US_Born" = 0,
                      "<5 years" = 1, 
                      "5-14 years" = 2 ,
                      "15+ years" = 3 ,
                      "N/A" = -15
                    ) ,
            label = "Years in US"
            )
        )

  assign(fyc_names[i], data)

}

```



### Avoidable Hospitalization 

```{r}

source(paste0(wd$codes, "avoid_hosptl.R"))


yearset <- 2011:2019
condsets <- vector("list", length(yearset))
fyc_names <- paste0("fyc", 11:19)  # Creates "fyc11", "fyc12", ..., "fyc15"
fycsets <- mget(fyc_names)
# Define ICD codes
icd10_codes <- list(
    hypertension = "I10",
    copd = c("J44", "J45"),
    diabetes = "E11",
    chf = "I50",
    pneumonia = "J18",
    uti = "N39"
)
icd9_codes <- list(
    diabetes = "250",
    hypertension = c("401", "402", "403"),
    copd = c("491", "490", "496", "466", "492", "494"),
    chf = c("428", "398"),
    pneumonia = c("480", "481", "482", "483", "486"),
    uti = "599"
)


# Loop through years and process

for (i in seq_along(yearset)) {
    condsets[[i]] <- read_MEPS(year = yearset[i], type = "COND")

    if (yearset[i] < 2016) {
        fycsets[[i]] <- process_condsets(condsets[[i]], fycsets[[i]], "ICD9CODX", icd9_codes)
    } else {
        fycsets[[i]] <- process_condsets(condsets[[i]], fycsets[[i]], "ICD10CDX", icd10_codes)
    }
      assign(fyc_names[i], fycsets[[i]])

           
}



```






### RACE

for year 2011, merge ethnicity into the race

```{r}

fyc11$RACETHX<-fyc2011$RACETHX
fyc11$RACETHX<-replace(fyc11$RACETHX,fyc11$RACETHX %in% c(6,3,5) ,5)  # Other Race
fyc11$RACETHX<-replace(fyc11$RACETHX,fyc11$RACETHX==2,3)  # Black only
fyc11$RACETHX<-replace(fyc11$RACETHX,fyc11$RACETHX %in% c(1), 2 ) # White 
fyc11$RACETHX<-ifelse(fyc11$HISPANX==1,1,fyc11$RACETHX) # Hispanic 


```

 fix the labels for race 

```{r}

fyc_names<- paste0("fyc", 11:19)
fycsets <- mget(fyc_names)
  
for (i in seq_along(fycsets)) {
 data <-fycsets[[i]]
 data <- data %>% 
    mutate (
        RACETHX = labelled(
            RACETHX, 
            labels = c(         
              "Hispanic" = 1,
              "Non-Hispanic White Only" = 2,
              "Non-Hispanic Black Only" = 3,
              "Non-Hispanic Asian Only" = 4,
              "Other Races" = 5
                    ) ,
            label = "Race & Ethnicity"
            )
        )
  assign(fyc_names[i], data)

}

```




### impute ADRISK42 and ADOVER42, ADSMOK42 brininging info from 2019 and 2017 
 for year 2018
```{r}

for (var in c("ADRISK42", "ADOVER42", "ADSMOK42")) {
  
    fyc18[[var]] <- NA

matching_indices<- match(fyc18$DUPERSID, fyc19$DUPERSID)
replacement <- fyc19[[var]][matching_indices[!is.na(matching_indices)]]
fyc18[[var]] <- replace(fyc18[[var]], !is.na(matching_indices)&fyc18$PANEL==23, replacement)


matching_indices <- match(fyc18$DUPERSID, paste0(as.character(fyc2017$PANEL), fyc2017$DUPERSID))
  replacement <- fyc2017[[var]][matching_indices[!is.na(matching_indices)]]
  fyc18[[var]] <- replace(fyc18[[var]], !is.na(matching_indices) & fyc18$PANEL == 22, replacement)
       
}

 fyc18 <- na.omit(fyc18)


```





## 5. Merge 2011 to 2019 in one data set

```{r}



allyear = bind_rows(fyc11,fyc12,fyc13,fyc14,fyc15,fyc16,fyc17,fyc18,fyc19) %>%
  mutate(poolwt = PERWT / 9)

#partyear = bind_rows(fyc11,fyc12,fyc13,fyc14,fyc15,fyc16,fyc18) %>%
  # (poolwt = PERWT / 7)

data <- allyear

```


## 6. Further cleaning

### Female

```{r}

data <- data %>%
  rename(female = SEX) %>%
  mutate(
   female = replace(female,female==1,0),
   female = replace(female,female==2,1) ,
    female = labelled(female, label = "Female")
)

table(data$female)

```

### Married 
```{r}

data$MARRYX <- replace(data$MARRYX,data$MARRYX %in% c(2,3,4,5),0)
data$MARRYX <- replace(data$MARRYX,data$MARRYX %in% c(-7,-8,-9),-15)
data$MARRYX <- labelled(data$MARRYX, label= "Married")

```

### Employed

```{r}
data$EMPST31 <- replace(data$EMPST31, data$EMPST31 %in% c(2,3,1),1) # Employed
data$EMPST31 <- replace(data$EMPST31, data$EMPST31 %in% c(-1,4),0) #unemployed
data$EMPST31 <- replace(data$EMPST31,data$EMPST31 %in% c(-7,-8,-9),-15)

data$EMPST31 <- labelled(data$EMPST31, label= "Employed")

```


### Health status: Good health 
```{r}

data$RTHLTH42 <- replace(data$RTHLTH42, data$RTHLTH42 %in% c(1,2,3),1) # Agree
data$RTHLTH42 <- replace(data$RTHLTH42, data$RTHLTH42 %in% c(4,5),0) # Disagree
data$RTHLTH42 <- replace(data$RTHLTH42, data$RTHLTH42 %in% c(-7,-8,-9,-15),-15) # Disagree

data$RTHLTH42 <- labelled(data$RTHLTH42, label = "Self Reported Good Health")

```

# Have US Provider

```{r}
data$HAVEUS42 <- replace(data$HAVEUS42, data$HAVEUS42 %in% c(-1),2) # Agree

```

### Risk taking

make it a dummy variable if person somewhat or strongly agrees that they are MORE LIKELY TO TAKE RISKS I recode this to 1, yes and if the person disagrees or uncertain recode to 2 or disagree

```{r}

data$ADRISK42 <- replace(data$ADRISK42, data$ADRISK42 %in% c(1,2,3),0) # Disagree
data$ADRISK42 <- replace(data$ADRISK42, data$ADRISK42 %in% c(4,5),1) # Agree
data$ADRISK42 <- replace(data$ADRISK42, data$ADRISK42 %in% c(-7,-8,-9,-15),-15) # Disagree

data$ADRISK42 <- labelled(data$ADRISK42, label = "Likely to Take Risk")
```

### overcome illnes on their own
```{r}

data$ADOVER42 <- replace(data$ADOVER42, data$ADOVER42 %in% c(1,2,3),0) # Disagree 
data$ADOVER42 <- replace(data$ADOVER42, data$ADOVER42 %in% c(4,5),1) # Agree
data$ADOVER42 <- replace(data$ADOVER42, data$ADOVER42 %in% c(-7,-8,-9,-15),-15) # Disagree

data$ADOVER42 <- labelled(data$ADOVER42, label = "Belief in Overcoming Illness Alone")

```

## HRQOL SVS12


```{r}
#mental components

data$MCS42 <- replace(data$MCS42, data$MCS42 %in% c(-7,-8,-9,-15), -15)

data$PCS42 <- replace(data$PCS42, data$PCS42 %in% c(-7,-8,-9,-15), -15)

```



## 7. Create variables

### fake state variable:

```{r }

# Load necessary library
set.seed(123)  # Setting seed for reproducibility


# Define the regions and corresponding states with FIPS codes
region_to_states <- list(
  "1" = c("09", "23", "25", "33", "34", "36", "42", "44", "50"),
  "2" = c("18", "17", "19", "20", "26", "27", "29", "31", "38","39","46", "55"),
  "3" = c("01", "05", "10", "11", "12", "13", "21", "22", "24","28", "37", "40", "45", "47", "51", "54", "48", "51", "54"),
  "4" = c("02", "04", "06", "08", "15", "16", "30", "32", "35","41", "49","53" ,"56")
)

# Function to randomly assign a state based on region
random_state <- function(region) {
  states <- region_to_states[[region]]
  if (is.null(states)) return(NA)
  sample(states, 1)
}

# note I need to drop inapplicable before using this variable
# 
data<-data[data$REGION != -1,] 
# Apply the function to create the state variable
data$state <- sapply(data$REGION, random_state)

# Print the updated data frame

data$state<-as.numeric(data$state)
```


### expansion variable
this is the code I need to give them and get the data set 

```{r}
data$Adopted<- ifelse (data$state %in% c(1,12,13,16, 20,23,28,29,31,37,40,45,46,47,48,49,51,55,56), 0, 1)

data$ExpansionY <- ifelse (data$state %in% c(4,5,6,8,9,10,11, 15,17,19, 21,24,
                                             25,26,27,32,34,35,36,38,39,41,44,50,53,54), 2014, 0)  # codes the state that didn't adopt medicaid 
data$ExpansionY <- ifelse (data$state %in% c(18,33,42), 2015,data$ExpansionY) # codes the state that didn't adopt medicaid 
data$ExpansionY <- ifelse (data$state %in% c(2,22,30), 2016, data$ExpansionY) # codes the state that didn't adopt medicaid 

 b<-ifelse (data$state %in% c(18,33,42), 2015,0)
table(b)
table(data$ExpansionY)
table(data$Adopted)

table(data$state, data$Adopted)
table(data$ExpansionY)

#data$treat <- ifelse(data$YEAR>=data$ExpansionY & !data$ExpansionY==0 ,1,0)

data$treat <- ifelse(data$YEAR>=data$ExpansionY & !data$ExpansionY==0 ,1,
                     ifelse(data$ExpansionY==0,999,0))

table(data$treat)


```


## 8. Import other variables

### State unemployment rate 
```{r,eval=TRUE}

UNEMPR <- read_excel(paste0(wd$data,"UnempR.xlsx"), sheet = "Sheet1")
sapply(UNEMPR, class)

# Drop the columns I don't want, Statae name and State ABR are not included in the original Dataset, 
# I might need to drop them or just keep the ABR for the final dataset

UNEMPR <- UNEMPR %>% select(-c(`STATE ABR` , `2020`, `2021`))

# Use pivot_longer to reshape the data
unemp_data_long <- pivot_longer(UNEMPR, 
                                cols = c(`2009`,`2010`, `2011`, `2012`, `2013`, `2014`, `2015`, `2016`, `2017`, `2018`, `2019`), 
                                names_to = "YEAR", 
                                values_to = "UnempR")

# Print the transformed data check if everything is correct
print(unemp_data_long)
UNEMPR<-unemp_data_long 
rm(unemp_data_long)

colnames(UNEMPR) <- c("StateN","state", "YEAR", "UnempR")
 #cheking the values again, year is charachter and need to be converted to numeric
sapply(UNEMPR, class)
UNEMPR$YEAR<- as.numeric(UNEMPR$YEAR)
table(UNEMPR$YEAR)

data <- merge(data, UNEMPR, by = c( "state", "YEAR" ), all.x = TRUE)

table(data$StateN)

table(data$state)

```


### State structural xenphobia

```{r,eval=TRUE}


IPC <- read.csv(paste0(wd$data,"IPC.csv"), header = TRUE, sep ="," , fill = TRUE)
#IPC <- read_excel(paste0(wd$data,"IPC.csv"), sheet = "Sheet1")

sapply(IPC, class)
colnames(IPC) <- c("StateN", "YEAR", "IPC","state")
IPC<- subset(IPC, select = -StateN)
data <- merge(data, IPC, by = c("state", "YEAR"), all.x = TRUE)

# fixing the missing values 
table(data$StateN, data$state)

table(data$StateN,data$IPC,exclude=NaN)
table(data$YEAR[data$StateN == "North Carolina"])
data$IPC<-ifelse(data$StateN == "North Carolina" & data$YEAR==2016 & is.na(data$IPC), -3, data$IPC)


```



## 9. Handeling Missing variables

```{r}

data[data == -15 | data == -9 | data== -8 | data==-7] <- -15

data[data == -15] <- NA


```

## drop extra variables"


```{r}

data <- data %>%
  select(-CABREAST , -CACOLON )

rm(fyc2011,fyc2012,fyc2013,fyc2014,fyc2015,fyc2016,fyc2017,fyc2018,fyc2019,fyc2020)
rm(fyc11,fyc12,fyc13,fyc14,fyc15,fyc16,fyc17,fyc18,fyc19)

```

## subset to older than 25


```{r}
data<- data %>% filter(AGE >25)

```




## 10. Save data for later

# RAW<-data


```{r}

 RAW<-data
#data<-RAW

```


```{r}


#save(data, file =paste0(wd$data,"RAWMEPS.RData"))

#load(paste0(wd$data,"RAWMEPS.RData"))

#MEPS<- data


#library(haven)

# Example: Save as Stata file
#write_dta(data, paste0(wd$data,"MEPSFAKE.dta"))


```


